---
title: "Common orthogroups"
author: "cagood"
date: "6.15.18"
output: html_notebook
---

Processing data from my 6.13.2018 OrthoFinder run (located in /home/cagood/Dimensions/OrthoFinder/pepfiles/Results_Jun13/) in order to return a list of orthogroups common to all 8 species.

```{r Read in Orthogroup count data}
genecounts = data.frame(read.table("/home/cagood/Dimensions/OrthoFinder/pepfiles/Results_Jun13/Orthogroups.GeneCount.csv", header = T))

head(genecounts)
```

```{r Subset}
# Add to subset if columns 1-8 are all greater than zero
# MARGIN = 1 sets apply to 'rows', so that each row's values must all be greater than 0
commonOGs = genecounts[apply(genecounts, MARGIN = 1, function(x) all(x > 0)), ]

head(commonOGs)
tail(commonOGs)
```

My commonOGs subset has 3156 observations, which matches the value in ~/Dimensions/OrthoFinder/pepfiles/Results_Jun13/Statistics_Overall.csv for 'Number of orthogroups with all species present'. Nice. 

```{r List of common OGs}
# Get rownames of commonOGs and output a list of these names
namelist = row.names(commonOGs)

write.table(namelist, 'commonOGs.txt')

head(namelist)
```

```{r Read in Orthogroup ID data}
# Read in the total list of genes assigned to Orthogroups
geneIDs = read.csv("/home/cagood/Dimensions/OrthoFinder/pepfiles/Results_Jun13/Orthogroups.csv", header = F, sep = '\t')

head(geneIDs)
```

```{r Filter geneIDs by commonOGs list}
# Subset the list of genes by orthogroup, based on whether the orthogroup is in the namelist, and write a new csv file
common_genes_by_OG = geneIDs[geneIDs$V1 %in% namelist,]

write.table(common_genes_by_OG, 'commonOG_Nodes.tsv', quote = F, sep = '\t', col.names = NA)
```

