---
title: "DEseq2: exploring, part ii"
author: Charles Goodman
output: html_notebook
---

Having worked through the basics of DESeq2 in part i, here I'll make an effort to tighten up the pipeline to be used for single-species analysis.

Input:  A Rosetta file
Output: A .csv of differentially (and constitutively) expressed genes for each consecutive pair of timepoints, and associated figures
Next:   Use the .csv to filter the original Rosetta for gene IDs, and write a script to pull GO terms for each

**Manuals & Pages**
[DEseq Homepage](https://bioconductor.org/packages/release/bioc/html/DESeq2.html)
[Sample Workthrough](https://bioconductor.org/packages/release/bioc/vignettes/DESeq2/inst/doc/DESeq2.html)
[Full Manual](https://bioconductor.org/packages/release/bioc/manuals/DESeq2/man/DESeq2.pdf)

## Packages and libaries
```{r include=FALSE}
# Initializing DEseq
source("https://bioconductor.org/biocLite.R")
biocLite("DESeq2")

# Packages - if unavailable use biocLite() from terminal in sudo R session
library(dplyr)
library(tibble)
library(DESeq2)
library(ggplot2)
library(BiocParallel)

# Parallel operation
register(MulticoreParam(8))
```

## Data initialization
```{r}
# Read in DC10_Rosetta.csv
ros_file = read.csv(file = "/home/cagood/Dimensions/R_analyses/New_Rosettas/Outputs/DC10_Rosetta.csv", header = T)

# Subset Rosetta data
# Use ID column for names, and .ec columns for counts
DC10_counts = select(ros_file,
                     t0.r1.ec, t0.r2.ec, t0.r3.ec, 
                     t3.r1.ec, t3.r2.ec, t3.r3.ec,
                     t5.r1.ec, t5.r2.ec, t5.r3.ec)

# Set ros_file ID column as list row_names
row_names = ros_file$ID

# Set count matrix values as integers (effectively, rounds down to nearest whole#)
DC10_counts = sapply(DC10_counts[,1:9], as.integer)

# Set count matrix row names = row_names
rownames(DC10_counts) = row_names

# Subset for each pair of consecutive timepoints
DC10_T1T3 = DC10_counts[,1:6]
DC10_T3T5 = DC10_counts[,4:9]
```

## Metadata initialization
```{r}
# Metadata for subsequent dds objects - prepare one for each 'counts' subset

# T1T3
coldat_T1T3 = data.frame(
  row.names = colnames(DC10_T1T3),
  condition = c("T1","T1","T1",
                "T3","T3","T3"),
  type = c("paired-end","paired-end","paired-end",
           "paired-end","paired-end","paired-end"))

# T3T5
coldat_T3T5 = data.frame(
  row.names = colnames(DC10_T3T5),
  condition = c("T3","T3","T3",
                "T5","T5","T5"),
  type = c("paired-end","paired-end","paired-end",
           "paired-end","paired-end","paired-end"))
```

## Generate dds objects
```{r}
# Prepare a dds for each 'counts' subset and associated metadatatable

# T1T3
dds_T1T3 <- DESeqDataSetFromMatrix(countData = DC10_T1T3,
                              colData = coldat_T1T3,
                              design = ~ condition)

# T3T5
dds_T3T5 <- DESeqDataSetFromMatrix(countData = DC10_T3T5,
                              colData = coldat_T3T5,
                              design = ~ condition)
```

## Prefiltering
```{r}
# Keep only dds rows with counts greater than 10

# T1T3
## keep = rowSums(counts(dds_T1T3)) >=10
## dds_T1T3 = dds_T1T3[keep,]

# T3T5
## keep = rowSums(counts(dds_T3T5)) >=10
## dds_T3T5 = dds_T3T5[keep,]

```

## DESeq analysis & pairwise comparison
```{r}
# T1T3
dds_T1T3 = DESeq(dds_T1T3)
res_T1T3 = results(dds_T1T3, contrast=c("condition","T3","T1")) # Note: contrast uses args (condition, numerator, denominator)
res_T1T3= res_T1T3[order(res_T1T3$pvalue),]

# T3T5
dds_T3T5 = DESeq(dds_T3T5)
res_T3T5 = results(dds_T3T5, contrast=c("condition","T5","T3")) # To set p value cutoff, set alpha = 0.05
res_T3T5= res_T3T5[order(res_T3T5$pvalue),]

# Show summary data
summary(res_T1T3)
summary(res_T3T5)
```

## plotMA()
Visualize LFC over mean of normalized counts - red = p<0.05
```{r}
plotMA(res_T1T3)
plotMA(res_T3T5)
```

## lfcShrink()
Shrinks effect size by removing noise associated with LFC from low count genes using apeglm (Zhu, A., Ibrahim, J.G., Love, M.I. (2018)

```{r}
# Check coefficient names for each dds, Pass coefficients from the above call into lfcShrink(), and plot. Output .csvs of results

# T1T3
resultsNames(dds_T1T3)
res_T1T3_LFC = lfcShrink(dds_T1T3, coef = "condition_T3_vs_T1", type = "apeglm")

# T3T5
resultsNames(dds_T3T5)
res_T3T5_LFC = lfcShrink(dds_T3T5, coef = "condition_T5_vs_T3", type = "apeglm")

# Plots
plotMA(res_T1T3_LFC, ylim=c(-10,10))
plotMA(res_T3T5_LFC, ylim=c(-10,10))
```

## .csvs and subsets
```{r}
# .csvs
write.csv(as.data.frame(res_T1T3), 
          file="./exploring_data/DC10_res_T1T3_results.csv")
write.csv(as.data.frame(res_T3T5), 
          file="./exploring_data/DC10_res_T3T5_results.csv")

# To subset, subset_eg = subset(res_eg, padj < 0.05)
```

## Rorshach plots
```{r}

sigP_T1T3_LFC = data.frame(subset(res_T1T3_LFC, padj < 0.05))
sigP_T3T5_LFC = data.frame(subset(res_T3T5_LFC, padj < 0.05))

commonSigPs = merge(sigP_T1T3_LFC$log2FoldChange, sigP_T3T5_LFC$log2FoldChange, by = "row.names")

plot(commonSigPs$x, commonSigPs$y, col = rgb(0,0,1,0.15), pch = 20)
lines(commonSigPs$x, commonSigPs$y, col = rgb(1,0,0,0.05), type = 'p')
lines(commonSigPs$x, commonSigPs$y, col = rgb(0,0,0,0.03), type = 'p')
```
I'm not sure this is better - it only shows genes differentially expressed in BOTH pairs of timepoints, which ignores a fair bit. 

Also, I've determined a significant issue here: DESeq2 doesn't play well with Kallisto data directly - I can't simply read in est. counts and expect things to work properly. I have 2 options:

1. Use tximport to work the kallisto data into an acceptable format
2. Use Pachter lab's 'sleuth' instead. That will likely be my MO. See https://www.biostars.org/p/143458/#157303