---
title: "Sleuth_exploring_monos.pairwise"
author: Charles Goodman
output: html_notebook
---

This notebook will be used to produce locally saved sleuth_objects and summary sleuth_tables for each of the monoculture datasets - except rather than a full timecourse analysis, we're considering consecutive pairs of timepoints. Here, exploratory run with DC10 - later I'll write a cleaner version with better varnames and comments.

*Intros and manuals*
```{r}
vignette('intro', package = 'sleuth')
help(package = 'sleuth')
```

Also see:

[GitHub](https://github.com/pachterlab/sleuth)
[Pachter Lab](https://pachterlab.github.io/sleuth/)

*Kallisto output data pooling*
```
# 1. Copied kallisto data from ~/Dimensions/R_analyses/Monoculture_Expression_Data to ~/Dimensions/Monocultures (orig. still intact!)
scp -r ./Data/ /home/cagood/Dimensions/Monocultures/

# 2. Renamed directory to mono_expr_data
mv Data/ mono_expr_data

# 3. Changed kallisto output directories from 'kallisto_t0_r1' etc. to 'DC10_t0_r1' etc... NOTE: do this for each species!
rename 's/kallisto/DC10/' kallisto_*/

# 4. Find and delete empty directories below current dir
find . -type d -empty -print
find . -type d -empty -delete

# Now each monoculture has specifically-named kallisto output data.
```

*Initialize packages*

```{r include=FALSE}
source("http://bioconductor.org/biocLite.R")
# Only need to install these packages once... do this from sudo R session in terminal
##biocLite("devtools")
##biocLite("pachterlab/sleuth")
##biocLite("rhdf5")
##biocLite("shinyapp")

library("dplyr")
library("sleuth")
library("rhdf5")
library("devtools")
library("shiny")
library("splines")
```

*Analyses, by pairwise timepoints*
Format a block for each pair of consecutive timepoints.

**Input** = kallisto data location, condition values, output var/filenames
**Output** = A local named sleuth_object, and a csv of sleuth_results() values

*Read in data*
```{r}
# Set sample_id to hold a list of directories containing kallisto data of interest
sample_id = dir("~/Dimensions/Monocultures/mono_expr_data/01")

# Collate a list of paths to kallisto results directories using sample_id
kal_dirs = file.path("~/Dimensions/Monocultures/mono_expr_data/01",sample_id)

# Construct a metadata table
metadata = data.frame(
  condition = c("T0","T0","T0",
                "T3","T3","T3",
                "T5","T5","T5"),
  sample = sample_id,
  path = as.character(kal_dirs))

metadata = data.frame(lapply(metadata, as.character), stringsAsFactors = FALSE)
```

*Construct a sleuth object*
```{r}
# Subset T1/T3 values
T0T3 = dplyr::select(metadata[1:6,], sample, condition, path)
T3T5 = dplyr::select(metadata[4:9,], sample, condition, path)

# Construct the sleuth object
so_T0T3 = sleuth_prep(T0T3)
so_T3T5 = sleuth_prep(T3T5)
```

*Fit models*
```{r}
so_T0T3 = sleuth_fit(so_T0T3, ~condition, 'full')
so_T0T3 = sleuth_fit(so_T0T3, ~1, 'reduced')

so_T3T5 = sleuth_fit(so_T3T5, ~condition, 'full')
so_T3T5 = sleuth_fit(so_T3T5, ~1, 'reduced')

# Check fit models
models(so_T0T3)
models(so_T3T5)
```

*Run Tests*
```{r}
so_T0T3 = sleuth_lrt(so_T0T3, 'reduced', 'full')
so_T3T5 = sleuth_lrt(so_T3T5, 'reduced', 'full')
```

*Get results and write tables*
```{r}
st_T0T3 = sleuth_results(so_T0T3, 'reduced:full', 'lrt', show_all = TRUE)
st_T3T5 = sleuth_results(so_T3T5, 'reduced:full', 'lrt', show_all = TRUE)

write.csv(st_T0T3, file = "./mono_sleuth_pairwise/DC10_T0T3_sleuth-tab", row.names = FALSE)
write.csv(st_T3T5, file = "./mono_sleuth_pairwise/DC10_T3T5_sleuth-tab", row.names = FALSE)
```

