---
title: "Sleuth_monos"
author: Charles Goodman
output: html_notebook
---

This notebook will be used to produce locally saved sleuth_objects and summary sleuth_tables for each of the monoculture datasets; these will ultimately serve a variety of purposes in downstream analysis. 

*Intros and manuals*
```{r}
vignette('intro', package = 'sleuth')
help(package = 'sleuth')
```

Also see:

[GitHub](https://github.com/pachterlab/sleuth)
[Pachter Lab](https://pachterlab.github.io/sleuth/)

*Kallisto output data pooling*
```
# 1. Copied kallisto data from ~/Dimensions/R_analyses/Monoculture_Expression_Data to ~/Dimensions/Monocultures (orig. still intact!)
scp -r ./Data/ /home/cagood/Dimensions/Monocultures/

# 2. Renamed directory to mono_expr_data
mv Data/ mono_expr_data

# 3. Changed kallisto output directories from 'kallisto_t0_r1' etc. to 'DC10_t0_r1' etc... NOTE: do this for each species!
rename 's/kallisto/DC10/' kallisto_*/

# 4. Find and delete empty directories below current dir
find . -type d -empty -print
find . -type d -empty -delete

# Now each monoculture has specifically-named kallisto output data.
```

*Initialize packages*

```{r include=FALSE}
source("http://bioconductor.org/biocLite.R")
# Only need to install these packages once... do this from sudo R session in terminal
##biocLite("devtools")
##biocLite("pachterlab/sleuth")
##biocLite("rhdf5")
##biocLite("shinyapp")

library("dplyr")
library("sleuth")
library("rhdf5")
library("devtools")
library("shiny")
library("splines")
```

*Analyses, by monoculture*
Format a block for each monoculture.

**Input** = kallisto data location, condition values, output var/filenames
**Output** = A local named sleuth_object, and a csv of sleuth_results() values

Notes: 

'condition' values reflect days past 1st RNA extraction, from DIMENSIONS_CompetitionMaster.xls. Set count @ first extraction in sample_id = 1
T0 = 14-Apr
T1 = 16-Apr
T2 = 18-Apr
T3 = 22-Apr
T4 = 24-Apr
T5 = 02-May
T6 = 22-May

df = degrees of freedom; set to lowest number in s2c matrix dimensions

```{r DC10, include=FALSE}
## DC10 Inputs
# Set sample_id to hold a list of directories containing kallisto data of interest
sample_id = dir("~/Dimensions/Monocultures/mono_expr_data/01")

# Collate a list of paths to kallisto results directories using sample_id
kal_dirs = file.path("~/Dimensions/Monocultures/mono_expr_data/01",sample_id)

# Construct a 'sample-to-covariate' table (I was calling this 'metadata' before)
# Set count @ first extraction in sample_id = 1
s2c = data.frame(
  sample = sample_id,
  condition = c(1,1,1,
                9,9,9,
                19,19,19),
  path = as.character(kal_dirs),
  stringsAsFactors = FALSE)

# Generate the model based on the s2c$condition values, w/ degrees of freedom = lowest number of matrix dimensions
design_matrix = model.matrix(formula(~ ns(s2c$condition, df=3)))


## Model fitting
# Prepare the sleuth object using the splines-based model
so <- sleuth_prep(s2c, full_model = design_matrix)

# Fit the full model, based on condition (day of RNA collection); and fit the reduced model w/o
so <- sleuth_fit(so, ~condition, 'full')
so <- sleuth_fit(so, formula = ~ 1, fit_name = "reduced")

# Perform likelihood ratio test
so <- sleuth_lrt(so, "reduced", "full")


## Outputs
# Set the sleuth_object and sleuth_table to a named local var
DC10so = so
DC10st = sleuth_results(so, 'reduced:full', 'lrt', show_all = TRUE)

# Write the sleuth_table to a .csv
write.csv(DC10st, file = "./mono_sleuth_tables/DC10_sleuth-tab", row.names = FALSE)
```

```{r DC20, include=FALSE}
## DC20 Inputs 
# Set sample_id to hold a list of directories containing kallisto data of interest
sample_id = dir("~/Dimensions/Monocultures/mono_expr_data/02")

# Collate a list of paths to kallisto results directories using sample_id
kal_dirs = file.path("~/Dimensions/Monocultures/mono_expr_data/02",sample_id)

# Construct a 'sample-to-covariate' table (I was calling this 'metadata' before)
# Set count @ first extraction in sample_id = 1
s2c = data.frame(
  sample = sample_id,
  condition = c(1,1,
                11,11,
                19,19,19),
  path = as.character(kal_dirs),
  stringsAsFactors = FALSE)

# Generate the model based on the s2c$condition values, w/ degrees of freedom = lowest number of matrix dimensions
design_matrix = model.matrix(formula(~ ns(s2c$condition, df=2)))


## Model fitting
# Prepare the sleuth object using the splines-based model
so <- sleuth_prep(s2c, full_model = design_matrix)

# Fit the full model, based on condition (day of RNA collection); and fit the reduced model w/o
so <- sleuth_fit(so, ~condition, 'full')
so <- sleuth_fit(so, formula = ~ 1, fit_name = "reduced")

# Perform likelihood ratio test
so <- sleuth_lrt(so, "reduced", "full")


## Outputs 
# Set the sleuth_object and sleuth_table to a named local var
DC20so = so
DC20st = sleuth_results(so, 'reduced:full', 'lrt', show_all = FALSE)

# Write the sleuth_table to a .csv
write.csv(DC20st, file = "./mono_sleuth_tables/DC20_sleuth-tab", row.names = FALSE)
```

```{r DC30, include=FALSE}
## DC30 Inputs 
# Set sample_id to hold a list of directories containing kallisto data of interest
sample_id = dir("~/Dimensions/Monocultures/mono_expr_data/03")

# Collate a list of paths to kallisto results directories using sample_id
kal_dirs = file.path("~/Dimensions/Monocultures/mono_expr_data/03",sample_id)

# Construct a 'sample-to-covariate' table (I was calling this 'metadata' before)
# Set count @ first extraction in sample_id = 1
s2c = data.frame(
  sample = sample_id,
  condition = c(1,1,1,
                9,9,9,
                19,19,19),
  path = as.character(kal_dirs),
  stringsAsFactors = FALSE)

# Generate the model based on the s2c$condition values, w/ degrees of freedom = lowest number of matrix dimensions
design_matrix = model.matrix(formula(~ ns(s2c$condition, df=3)))


## Model fitting
# Prepare the sleuth object using the splines-based model
so <- sleuth_prep(s2c, full_model = design_matrix)

# Fit the full model, based on condition (day of RNA collection); and fit the reduced model w/o
so <- sleuth_fit(so, ~condition, 'full')
so <- sleuth_fit(so, formula = ~ 1, fit_name = "reduced")

# Perform likelihood ratio test
so <- sleuth_lrt(so, "reduced", "full")


## Outputs 
# Set the sleuth_object and sleuth_table to a named local var
DC30so = so
DC30st = sleuth_results(so, 'reduced:full', 'lrt', show_all = FALSE)

# Write the sleuth_table to a .csv
write.csv(DC30st, file = "./mono_sleuth_tables/DC30_sleuth-tab", row.names = FALSE)
```

```{r DC40, include=FALSE}
## DC40 Inputs 
# Set sample_id to hold a list of directories containing kallisto data of interest
sample_id = dir("~/Dimensions/Monocultures/mono_expr_data/04")

# Collate a list of paths to kallisto results directories using sample_id
kal_dirs = file.path("~/Dimensions/Monocultures/mono_expr_data/04",sample_id)

# Construct a 'sample-to-covariate' table (I was calling this 'metadata' before)
# Set count @ first extraction in sample_id = 1
s2c = data.frame(
  sample = sample_id,
  condition = c(1,1,1,
                7,7,7),
  path = as.character(kal_dirs),
  stringsAsFactors = FALSE)

# Generate the model based on the s2c$condition values, w/ degrees of freedom = lowest number of matrix dimensions
design_matrix = model.matrix(formula(~ ns(s2c$condition, df=2)))


## Model fitting
# Prepare the sleuth object using the splines-based model
so <- sleuth_prep(s2c, full_model = design_matrix)

# Fit the full model, based on condition (day of RNA collection); and fit the reduced model w/o
so <- sleuth_fit(so, ~condition, 'full')
so <- sleuth_fit(so, formula = ~ 1, fit_name = "reduced")

# Perform likelihood ratio test
so <- sleuth_lrt(so, "reduced", "full")


## Outputs 
# Set the sleuth_object and sleuth_table to a named local var
DC40so = so
DC40st = sleuth_results(so, 'reduced:full', 'lrt', show_all = FALSE)

# Write the sleuth_table to a .csv
write.csv(DC40st, file = "./mono_sleuth_tables/DC40_sleuth-tab", row.names = FALSE)
```

```{r DC50, include=FALSE}
## DC50 Inputs 
# Set sample_id to hold a list of directories containing kallisto data of interest
sample_id = dir("~/Dimensions/Monocultures/mono_expr_data/05")

# Collate a list of paths to kallisto results directories using sample_id
kal_dirs = file.path("~/Dimensions/Monocultures/mono_expr_data/05",sample_id)

# Construct a 'sample-to-covariate' table (I was calling this 'metadata' before)
# Set count @ first extraction in sample_id = 1
s2c = data.frame(
  sample = sample_id,
  condition = c(1,1,
                4,4,4,
                16,16,16),
  path = as.character(kal_dirs),
  stringsAsFactors = FALSE)

# Generate the model based on the s2c$condition values, w/ degrees of freedom = lowest number of matrix dimensions
design_matrix = model.matrix(formula(~ ns(s2c$condition, df=2)))


## Model fitting
# Prepare the sleuth object using the splines-based model
so <- sleuth_prep(s2c, full_model = design_matrix)

# Fit the full model, based on condition (day of RNA collection); and fit the reduced model w/o
so <- sleuth_fit(so, ~condition, 'full')
so <- sleuth_fit(so, formula = ~ 1, fit_name = "reduced")

# Perform likelihood ratio test
so <- sleuth_lrt(so, "reduced", "full")


## Outputs 
# Set the sleuth_object and sleuth_table to a named local var
DC50so = so
DC50st = sleuth_results(so, 'reduced:full', 'lrt', show_all = FALSE)

# Write the sleuth_table to a .csv
write.csv(DC50st, file = "./mono_sleuth_tables/DC50_sleuth-tab", row.names = FALSE)
```

```{r DC60, include=FALSE}
## DC60 Inputs 
# Set sample_id to hold a list of directories containing kallisto data of interest
sample_id = dir("~/Dimensions/Monocultures/mono_expr_data/06")

# Collate a list of paths to kallisto results directories using sample_id
kal_dirs = file.path("~/Dimensions/Monocultures/mono_expr_data/06",sample_id)

# Construct a 'sample-to-covariate' table (I was calling this 'metadata' before)
# Set count @ first extraction in sample_id = 1
s2c = data.frame(
  sample = sample_id,
  condition = c(1,1,1,
                4,4,4,
                16,16,16),
  path = as.character(kal_dirs),
  stringsAsFactors = FALSE)

# Generate the model based on the s2c$condition values, w/ degrees of freedom = lowest number of matrix dimensions
design_matrix = model.matrix(formula(~ ns(s2c$condition, df=3)))


## Model fitting
# Prepare the sleuth object using the splines-based model
so <- sleuth_prep(s2c, full_model = design_matrix)

# Fit the full model, based on condition (day of RNA collection); and fit the reduced model w/o
so <- sleuth_fit(so, ~condition, 'full')
so <- sleuth_fit(so, formula = ~ 1, fit_name = "reduced")

# Perform likelihood ratio test
so <- sleuth_lrt(so, "reduced", "full")


## Outputs 
# Set the sleuth_object and sleuth_table to a named local var
DC60so = so
DC60st = sleuth_results(so, 'reduced:full', 'lrt', show_all = FALSE)

# Write the sleuth_table to a .csv
write.csv(DC60st, file = "./mono_sleuth_tables/DC60_sleuth-tab", row.names = FALSE)
```

```{r DC70, include=FALSE}
## DC70 Inputs 
# Set sample_id to hold a list of directories containing kallisto data of interest
sample_id = dir("~/Dimensions/Monocultures/mono_expr_data/07")

# Collate a list of paths to kallisto results directories using sample_id
kal_dirs = file.path("~/Dimensions/Monocultures/mono_expr_data/07",sample_id)

# Construct a 'sample-to-covariate' table (I was calling this 'metadata' before)
# Set count @ first extraction in sample_id = 1
s2c = data.frame(
  sample = sample_id,
  condition = c(1,1,1,
                11,11,11),
  path = as.character(kal_dirs),
  stringsAsFactors = FALSE)

# Generate the model based on the s2c$condition values, w/ degrees of freedom = lowest number of matrix dimensions
design_matrix = model.matrix(formula(~ ns(s2c$condition, df=2)))


## Model fitting
# Prepare the sleuth object using the splines-based model
so <- sleuth_prep(s2c, full_model = design_matrix)

# Fit the full model, based on condition (day of RNA collection); and fit the reduced model w/o
so <- sleuth_fit(so, ~condition, 'full')
so <- sleuth_fit(so, formula = ~ 1, fit_name = "reduced")

# Perform likelihood ratio test
so <- sleuth_lrt(so, "reduced", "full")


## Outputs 
# Set the sleuth_object and sleuth_table to a named local var
DC70so = so
DC70st = sleuth_results(so, 'reduced:full', 'lrt', show_all = FALSE)

# Write the sleuth_table to a .csv
write.csv(DC70st, file = "./mono_sleuth_tables/DC70_sleuth-tab", row.names = FALSE)
```

```{r DC80, include=FALSE}
## DC80 Inputs 
# Set sample_id to hold a list of directories containing kallisto data of interest
sample_id = dir("~/Dimensions/Monocultures/mono_expr_data/08")

# Collate a list of paths to kallisto results directories using sample_id
kal_dirs = file.path("~/Dimensions/Monocultures/mono_expr_data/08",sample_id)

# Construct a 'sample-to-covariate' table (I was calling this 'metadata' before)
# Set count @ first extraction in sample_id = 1
s2c = data.frame(
  sample = sample_id,
  condition = c(1,1,1,
                3,3,3,
                10,10,10),
  path = as.character(kal_dirs),
  stringsAsFactors = FALSE)

# Generate the model based on the s2c$condition values, w/ degrees of freedom = lowest number of matrix dimensions
design_matrix = model.matrix(formula(~ ns(s2c$condition, df=3)))


## Model fitting
# Prepare the sleuth object using the splines-based model
so <- sleuth_prep(s2c, full_model = design_matrix)

# Fit the full model, based on condition (day of RNA collection); and fit the reduced model w/o
so <- sleuth_fit(so, ~condition, 'full')
so <- sleuth_fit(so, formula = ~ 1, fit_name = "reduced")

# Perform likelihood ratio test
so <- sleuth_lrt(so, "reduced", "full")


## Outputs 
# Set the sleuth_object and sleuth_table to a named local var
DC80so = so
DC80st = sleuth_results(so, 'reduced:full', 'lrt', show_all = FALSE)

# Write the sleuth_table to a .csv
write.csv(DC80st, file = "./mono_sleuth_tables/DC80_sleuth-tab", row.names = FALSE)
```

*Exploratory analysis*

```{r}
# How many significant results w/ q<0.1?
table(DC10st[,"qval"] < 0.1)
# table(DC20st[,"qval"] < 0.1)
# table(DC30st[,"qval"] < 0.1)
# table(DC40st[,"qval"] < 0.1)
# table(DC50st[,"qval"] < 0.1)
# table(DC60st[,"qval"] < 0.1)
# table(DC70st[,"qval"] < 0.1)
# table(DC80st[,"qval"] < 0.1)
```

A few different plots (from the vignette)

```{r}
# 1
plot_qq(DC10so, test = 'reduced:full', test_type = 'lrt', sig_level = 0.1)
```

```{r}
# 2
lrt_results <- sleuth_results(so, 'reduced:full', test_type = 'lrt')
table(lrt_results[,"qval"] < 0.1)
```

```{r}
# 3
plot_transcript_heatmap(so, head(lrt_results, n = 20)$target_id, 'tpm')
```

```{r}
# 4
library(ggplot2)
tmp <- DC10so$obs_raw %>% dplyr::filter(target_id == 'NODE_788_length_5499_cov_347.191349_g462_i0')
tmp <- dplyr::full_join(DC10so$sample_to_covariates, tmp, by = 'sample')
tmp
tmp <- transform(tmp, condition = as.numeric(condition))

ggplot(tmp, aes(x=condition, y=tpm)) + geom_point(shape=1) + geom_smooth(method = loess)
```