---
title: "Stringent Timecourse monoculture Analyses"
author: Charles Goodman
output: html_notebook
---

This notebook will be used to produce locally saved sleuth_objects and summary sleuth_tables from full timecourse analysis for each of the monoculture datasets; these will ultimately serve a variety of purposes in downstream analysis. 

*Intros and manuals*
```{r}
vignette('intro', package = 'sleuth')
help(package = 'sleuth')
```

Also see:

[GitHub](https://github.com/pachterlab/sleuth)
[Pachter Lab](https://pachterlab.github.io/sleuth/)

*Initialize packages*

```{r include=FALSE}
source("http://bioconductor.org/biocLite.R")
# Only need to install these packages once... do this from sudo R session in terminal
##biocLite("devtools")
##biocLite("pachterlab/sleuth")
##biocLite("rhdf5")
##biocLite("shinyapp")

library("dplyr")
library("sleuth")
library("rhdf5")
library("devtools")
library("shiny")
library("splines")
```

*Analyses, by monoculture*
Format a block for each monoculture.

**Input** = kallisto data location, condition values, output var/filenames
**Output** = A local named sleuth_object, and a csv of sleuth_results() values

Notes: 

For Timecourse analysis 'condition', convert T# = date = absolute day

T0  = 14-Apr = 1 
T1  = 16-Apr = 3
T2  = 18-Apr = 5
T3  = 22-Apr = 9
T4  = 24-Apr = 11
T5  = 02-May = 18
TY  = 10-May = 27
TWB = 17-May = 33
T6  = 22-May = 38


df = degrees of freedom; set to lowest number in s2c matrix dimensions

```{r DC10, include=FALSE}
## DC10 Inputs
# Set sample_id to hold a list of directories containing kallisto data of interest
sample_id = dir("~/Dimensions/Kallistos/stringent/DC10/")

# Collate a list of paths to kallisto results directories using sample_id
kal_dirs = file.path("~/Dimensions/Kallistos/stringent/DC10",sample_id)

# Construct a 'sample-to-covariate' table (I was calling this 'metadata' before)
# Set count @ first extraction in sample_id = 1
s2c = data.frame(
  sample = sample_id,
  condition = c(1,9,18,38,
                1,9,18,38,
                1,9,18,38),
  path = as.character(kal_dirs),
  stringsAsFactors = FALSE)

# Generate the model based on the s2c$condition values, w/ degrees of freedom = lowest number of matrix dimensions
design_matrix = model.matrix(formula(~ ns(s2c$condition, df=3)))


## Model fitting
# Prepare the sleuth object using the splines-based model
so <- sleuth_prep(s2c, full_model = design_matrix)

# Fit the full model, based on condition (day of RNA collection); and fit the reduced model w/o
so <- sleuth_fit(so, ~condition, 'full')
so <- sleuth_fit(so, formula = ~ 1, fit_name = "reduced")

# Perform likelihood ratio test
so <- sleuth_lrt(so, "reduced", "full")


## Outputs
st = sleuth_results(so, 'reduced:full', 'lrt', show_all = TRUE)
table(st[,"qval"] < 0.1)

# Write the sleuth_table to a .csv
write.csv(st, file = "./timecourse_Outs/DC10_stringent_timecourse", row.names = FALSE)

# cleanup
design_matrix = NULL
s2c = NULL
so = NULL
st = NULL
```

```{r DC20, include=FALSE}
## DC20 Inputs
# Set sample_id to hold a list of directories containing kallisto data of interest
sample_id = dir("~/Dimensions/Kallistos/stringent/DC20/")

# Collate a list of paths to kallisto results directories using sample_id
kal_dirs = file.path("~/Dimensions/Kallistos/stringent/DC20",sample_id)

# Construct a 'sample-to-covariate' table (I was calling this 'metadata' before)
# Set count @ first extraction in sample_id = 1
s2c = data.frame(
  sample = sample_id,
  condition = c(3,11,19,38,
                3,11,19,38,
                19,38),
  path = as.character(kal_dirs),
  stringsAsFactors = FALSE)

# Generate the model based on the s2c$condition values, w/ degrees of freedom = lowest number of matrix dimensions
design_matrix = model.matrix(formula(~ ns(s2c$condition, df=3)))


## Model fitting
# Prepare the sleuth object using the splines-based model
so <- sleuth_prep(s2c, full_model = design_matrix)

# Fit the full model, based on condition (day of RNA collection); and fit the reduced model w/o
so <- sleuth_fit(so, ~condition, 'full')
so <- sleuth_fit(so, formula = ~ 1, fit_name = "reduced")

# Perform likelihood ratio test
so <- sleuth_lrt(so, "reduced", "full")


## Outputs
st = sleuth_results(so, 'reduced:full', 'lrt', show_all = TRUE)
table(st[,"qval"] < 0.1)

# Write the sleuth_table to a .csv
write.csv(st, file = "./timecourse_Outs/DC20_stringent_timecourse", row.names = FALSE)

# cleanup
design_matrix = NULL
s2c = NULL
so = NULL
st = NULL
```

```{r DC30, include=FALSE}
## DC30 Inputs
# Set sample_id to hold a list of directories containing kallisto data of interest
sample_id = dir("~/Dimensions/Kallistos/stringent/DC30/")

# Collate a list of paths to kallisto results directories using sample_id
kal_dirs = file.path("~/Dimensions/Kallistos/stringent/DC30",sample_id)

# Construct a 'sample-to-covariate' table (I was calling this 'metadata' before)
# Set count @ first extraction in sample_id = 1
s2c = data.frame(
  sample = sample_id,
  condition = c(3,9,19,38,
                3,9,19,38,
                3,9,19,38),
  path = as.character(kal_dirs),
  stringsAsFactors = FALSE)

# Generate the model based on the s2c$condition values, w/ degrees of freedom = lowest number of matrix dimensions
design_matrix = model.matrix(formula(~ ns(s2c$condition, df=3)))


## Model fitting
# Prepare the sleuth object using the splines-based model
so <- sleuth_prep(s2c, full_model = design_matrix)

# Fit the full model, based on condition (day of RNA collection); and fit the reduced model w/o
so <- sleuth_fit(so, ~condition, 'full')
so <- sleuth_fit(so, formula = ~ 1, fit_name = "reduced")

# Perform likelihood ratio test
so <- sleuth_lrt(so, "reduced", "full")


## Outputs
st = sleuth_results(so, 'reduced:full', 'lrt', show_all = TRUE)
table(st[,"qval"] < 0.1)

# Write the sleuth_table to a .csv
write.csv(st, file = "./timecourse_Outs/DC30_stringent_timecourse", row.names = FALSE)

# cleanup
design_matrix = NULL
s2c = NULL
so = NULL
st = NULL
```

```{r DC40, include=FALSE}
## DC40 Inputs 
# Set sample_id to hold a list of directories containing kallisto data of interest
sample_id = dir("~/Dimensions/Kallistos/stringent/DC40/")

# Collate a list of paths to kallisto results directories using sample_id
kal_dirs = file.path("~/Dimensions/Kallistos/stringent/DC40",sample_id)

# Construct a 'sample-to-covariate' table (I was calling this 'metadata' before)
# Set count @ first extraction in sample_id = 1
s2c = data.frame(
  sample = sample_id,
  condition = c(5,11,38,27,        # NOTE: converted 'TY' to '27'
                5,11,38,27,
                5,11,38,27),
  path = as.character(kal_dirs),
  stringsAsFactors = FALSE)

# Generate the model based on the s2c$condition values, w/ degrees of freedom = lowest number of matrix dimensions
design_matrix = model.matrix(formula(~ ns(s2c$condition, df=3)))


## Model fitting
# Prepare the sleuth object using the splines-based model
so <- sleuth_prep(s2c, full_model = design_matrix)

# Fit the full model, based on condition (day of RNA collection); and fit the reduced model w/o
so <- sleuth_fit(so, ~condition, 'full')
so <- sleuth_fit(so, formula = ~ 1, fit_name = "reduced")

# Perform likelihood ratio test
so <- sleuth_lrt(so, "reduced", "full")


## Outputs 
st = sleuth_results(so, 'reduced:full', 'lrt', show_all = TRUE)
table(st[,"qval"] < 0.1)

# Write the sleuth_table to a .csv
write.csv(st, file = "./timecourse_Outs/DC40_stringent_timecourse", row.names = FALSE)

# cleanup
design_matrix = NULL
s2c = NULL
so = NULL
st = NULL
```

```{r DC50, include=FALSE}
## DC50 Inputs 
# Set sample_id to hold a list of directories containing kallisto data of interest
sample_id = dir("~/Dimensions/Kallistos/stringent/DC50/")

# Collate a list of paths to kallisto results directories using sample_id
kal_dirs = file.path("~/Dimensions/Kallistos/stringent/DC50",sample_id)

# Construct a 'sample-to-covariate' table (I was calling this 'metadata' before)
# Set count @ first extraction in sample_id = 1
s2c = data.frame(
  sample = sample_id,
  condition = c(9,19,33,        # NOTE: converted 'TWB' to '33'
                5,9,19,33,
                5,9,19,33),
  path = as.character(kal_dirs),
  stringsAsFactors = FALSE)

# Generate the model based on the s2c$condition values, w/ degrees of freedom = lowest number of matrix dimensions
design_matrix = model.matrix(formula(~ ns(s2c$condition, df=3)))


## Model fitting
# Prepare the sleuth object using the splines-based model
so <- sleuth_prep(s2c, full_model = design_matrix)

# Fit the full model, based on condition (day of RNA collection); and fit the reduced model w/o
so <- sleuth_fit(so, ~condition, 'full')
so <- sleuth_fit(so, formula = ~ 1, fit_name = "reduced")

# Perform likelihood ratio test
so <- sleuth_lrt(so, "reduced", "full")


## Outputs 
st = sleuth_results(so, 'reduced:full', 'lrt', show_all = TRUE)
table(st[,"qval"] < 0.1)

# Write the sleuth_table to a .csv
write.csv(st, file = "./timecourse_Outs/DC50_stringent_timecourse", row.names = FALSE)

# cleanup
design_matrix = NULL
s2c = NULL
so = NULL
st = NULL
```

```{r DC60, include=FALSE}
## DC60 Inputs 
# Set sample_id to hold a list of directories containing kallisto data of interest
sample_id = dir("~/Dimensions/Kallistos/stringent/DC60/")

# Collate a list of paths to kallisto results directories using sample_id
kal_dirs = file.path("~/Dimensions/Kallistos/stringent/DC60",sample_id)

# Construct a 'sample-to-covariate' table (I was calling this 'metadata' before)
# Set count @ first extraction in sample_id = 1
s2c = data.frame(
  sample = sample_id,
  condition = c(5,9,19,38,        # NOTE: converted 'TY' to '27'
                5,9,19,38,
                5,9,19,38),
  path = as.character(kal_dirs),
  stringsAsFactors = FALSE)

# Generate the model based on the s2c$condition values, w/ degrees of freedom = lowest number of matrix dimensions
design_matrix = model.matrix(formula(~ ns(s2c$condition, df=3)))


## Model fitting
# Prepare the sleuth object using the splines-based model
so <- sleuth_prep(s2c, full_model = design_matrix)

# Fit the full model, based on condition (day of RNA collection); and fit the reduced model w/o
so <- sleuth_fit(so, ~condition, 'full')
so <- sleuth_fit(so, formula = ~ 1, fit_name = "reduced")

# Perform likelihood ratio test
so <- sleuth_lrt(so, "reduced", "full")


## Outputs 
st = sleuth_results(so, 'reduced:full', 'lrt', show_all = TRUE)
table(st[,"qval"] < 0.1)

# Write the sleuth_table to a .csv
write.csv(st, file = "./timecourse_Outs/DC60_stringent_timecourse", row.names = FALSE)

# cleanup
design_matrix = NULL
s2c = NULL
so = NULL
st = NULL
```

```{r DC70, include=FALSE}
## DC70 Inputs 
# Set sample_id to hold a list of directories containing kallisto data of interest
sample_id = dir("~/Dimensions/Kallistos/stringent/DC70/")

# Collate a list of paths to kallisto results directories using sample_id
kal_dirs = file.path("~/Dimensions/Kallistos/stringent/DC70",sample_id)

# Construct a 'sample-to-covariate' table (I was calling this 'metadata' before)
# Set count @ first extraction in sample_id = 1
s2c = data.frame(
  sample = sample_id,
  condition = c(3,11,33,        # NOTE: converted 'TY' to '27' and 'TWB' to '33' - this one may have weird results
                3,11,38,27,
                3,11,33,27),
  path = as.character(kal_dirs),
  stringsAsFactors = FALSE)

# Generate the model based on the s2c$condition values, w/ degrees of freedom = lowest number of matrix dimensions
design_matrix = model.matrix(formula(~ ns(s2c$condition, df=3)))


## Model fitting
# Prepare the sleuth object using the splines-based model
so <- sleuth_prep(s2c, full_model = design_matrix)

# Fit the full model, based on condition (day of RNA collection); and fit the reduced model w/o
so <- sleuth_fit(so, ~condition, 'full')
so <- sleuth_fit(so, formula = ~ 1, fit_name = "reduced")

# Perform likelihood ratio test
so <- sleuth_lrt(so, "reduced", "full")


## Outputs 
st = sleuth_results(so, 'reduced:full', 'lrt', show_all = TRUE)
table(st[,"qval"] < 0.1)

# Write the sleuth_table to a .csv
write.csv(st, file = "./timecourse_Outs/DC70_stringent_timecourse", row.names = FALSE)

# cleanup
design_matrix = NULL
s2c = NULL
so = NULL
st = NULL
```

```{r DC80, include=FALSE}
## DC80 Inputs 
# Set sample_id to hold a list of directories containing kallisto data of interest
sample_id = dir("~/Dimensions/Kallistos/stringent/DC80/")

# Collate a list of paths to kallisto results directories using sample_id
kal_dirs = file.path("~/Dimensions/Kallistos/stringent/DC80",sample_id)

# Construct a 'sample-to-covariate' table (I was calling this 'metadata' before)
# Set count @ first extraction in sample_id = 1
s2c = data.frame(
  sample = sample_id,
  condition = c(3,5,11,38,        # NOTE: converted 'TY' to '27'
                3,5,11,38,
                3,5,11,38),
  path = as.character(kal_dirs),
  stringsAsFactors = FALSE)

# Generate the model based on the s2c$condition values, w/ degrees of freedom = lowest number of matrix dimensions
design_matrix = model.matrix(formula(~ ns(s2c$condition, df=3)))


## Model fitting
# Prepare the sleuth object using the splines-based model
so <- sleuth_prep(s2c, full_model = design_matrix)

# Fit the full model, based on condition (day of RNA collection); and fit the reduced model w/o
so <- sleuth_fit(so, ~condition, 'full')
so <- sleuth_fit(so, formula = ~ 1, fit_name = "reduced")

# Perform likelihood ratio test
so <- sleuth_lrt(so, "reduced", "full")


## Outputs 
st = sleuth_results(so, 'reduced:full', 'lrt', show_all = TRUE)
table(st[,"qval"] < 0.1)

# Write the sleuth_table to a .csv
write.csv(st, file = "./timecourse_Outs/DC80_stringent_timecourse", row.names = FALSE)

# cleanup
design_matrix = NULL
s2c = NULL
so = NULL
st = NULL
```
