---
title: "R Notebook"
output: html_notebook
---

I want to employ upset plots (as suggested by SM) in leu of venn diagrams. To do so, I'll try [UpSetR](https://caleydo.org/tools/upset/); [GitHub](https://github.com/hms-dbmi/UpSetR).

```{r}
library(UpSetR)
```

There are a lot of options as to what I can plot, so I'll start with a proof-of-concept: showing orthology across all 8 species, in terms of presence of orthogroups. 

```{r}
# Plotting OrthoGroup overlap

# Get a list of unique orthogroups for each species (ignoring repeats/isotypes)
species = c("DC10","DC20","DC30","DC40","DC50","DC60","DC70","DC80")

for (i in 1:8){
  name = species[i]
  
  path = paste0("/home/cagood/Dimensions/R_analyses/New_Rosettas/Outputs_mono/",species[i],"_Rosetta.csv")
  rosetta = read.csv(file = path)
  
  out = unique(rosetta$OG)
  assign(name, out)
}

# Make a list of those lists
ListInput = list(C.sorokiniana = DC10, 
                 C.acicularis = DC20,
                 C.turpinii = DC30,
                 P.charkowiensis = DC40,
                 S.acuminatus = DC50,
                 R.subcapitata = DC60,
                 S.puntulatum = DC70,
                 T.minimum = DC80)
```

```{r}
# List input
## png("upset.png")

upset(fromList(ListInput), 
      nsets = 8, 
      order.by = "freq", 
      number.angles = 40,
      mainbar.y.label = "Common Orthogroup Intersection",
      sets.x.label = "Orthogroups per species",
      query.legend = "top",
      queries = list(
        list(query = intersects, params = list("C.turpinii","S.puntulatum"), color = "orange", active = T, query.name = "Charophyceae"),
        list(query = intersects, params = list("C.sorokiniana","C.acicularis","P.charkowiensis","S.acuminatus","R.subcapitata","T.minimum"), color = "blue", active = T, query.name = "Chlorophyceae"))
)

## dev.off()

```


***


```{r}

# A contingency table might be more easily manipulated?

# Loop that shit in
species = c("DC10","DC20","DC30","DC40","DC50","DC60","DC70","DC80")
fullnames = c("C.sorokiniana", "C.acicularis", "C.turpinii" ,"P.charkowiensis" ,"S.acuminatus" ,"R.subcapitata" ,"S.puntulatum" ,"T.minimum" )

for (i in 1:8){
  name = species[i]
  
  path = paste0("/home/cagood/Dimensions/R_analyses/New_Rosettas/Outputs_mono/",species[i],"_Rosetta.csv")
  rosetta = read.csv(file = path)
  
  uniqueOGs = unique(rosetta$OG) # Right now ignore metadata, just use Unique OGs
  df = data.frame(matrix(unlist(uniqueOGs), nrow=length(uniqueOGs), byrow=T))
  colnames(df)[1] = "OG"
  df = cbind(spec = fullnames[i], df)
  out = df
  assign(name, out)
}

longdf = rbind(DC10,DC20,DC30,DC40,DC50,DC60,DC70,DC80)
TableInput = data.frame(t(as.data.frame.matrix(table(longdf)))) # Kinda wacky, but it does the thing. 
TableInput <- cbind(OGs = rownames(TableInput), TableInput)
row.names(TableInput) = NULL
```


Using a table for input makes it more easy to manipulate the resulting plots, can include metadata etc... Play with this later. 

```{r}
# Table Input
pdf("upset2.pdf")

upset(TableInput, 
      sets = c( "C.turpinii","S.puntulatum","C.sorokiniana","C.acicularis","P.charkowiensis","R.subcapitata" ,"S.acuminatus" ,"T.minimum"), #in phylo order
      keep.order = T,
      order.by = "freq",
      mainbar.y.label = "Common Orthogroup Intersection",
      sets.x.label = "Orthogroups per species",
      query.legend = "top",
      queries = list(
        list(query = intersects, params = list("C.turpinii","S.puntulatum"), color = "orange", active = T, query.name = "Charophyceae"),
        list(query = intersects, params = list("C.sorokiniana", "C.acicularis", "P.charkowiensis" ,"S.acuminatus" ,"R.subcapitata" ,"T.minimum" ), color = "blue", active = T, query.name = "Chlorophyceae"))
)

dev.off()

upset(TableInput, nsets = 8) 
```



